---
title: "Where Do They Buy It? Private-Label Performance by Department"
subtitle: "BZAN 535 | Team Project 1 | Individual section: Ximin Zeng"
author: "Ximin Zeng (xzeng3@utk.edu)"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    code_folding: show
params:
  trans_path: "C:/transaction_data.csv"
  prod_path:  "C:/product.csv"
  demo_path:  "C:/hh_demographic.csv"
  min_hh_dept: 30           # household must spend >= $30 in a dept to be included
  n_focus: 5                # how many departments carried into the tests
  boot_reps: 1000           # cluster-bootstrap replications
  seed: 535
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.width = 8, fig.height = 4.8, dpi = 300, fig.align = "center"
)
options(scipen = 999, dplyr.summarise.inform = FALSE)
set.seed(params$seed)
```

# Purpose and scope of this section

The team question is *"Who buys our private-label products, where do they buy them,
and what should we do with that information?"* My piece is the **where — which
part of the store**:

> **In which departments does Kroger's private label actually win, and are the
> differences between departments large enough and certain enough to act on?**

Two deliberately different things are measured, because merchandising decisions
need both:

| Question | Metric | Why it matters |
|---|---|---|
| Where is the private-label **business** today? | Private-label **revenue** (absolute $) | Sizes the prize; where a point of share is worth the most money |
| Where does private label **win** with shoppers? | Private-label **dollar share** within department | Removes department size so a small aisle can still look strong |

These rank departments differently. Reporting only one is the most common
mistake in this kind of analysis, so both appear side by side.

## Unit of analysis (stated up front, applied throughout)

1. **Line item (one product in one basket)** — used *only* to total revenue.
   Revenue is a sum of dollars; the line item is the accounting atom.
2. **Household × department** — the **primary unit for all inference.** One
   observation is *one household's private-label dollar share inside one
   department*. The decision is about shopper behaviour, not transactions: a
   single household making 400 trips would otherwise dominate any item- or
   basket-level test. Aggregating to the household gives every shopper one vote.
3. **Product (one row of `product.csv`)** — used only in the robustness section,
   to measure how much private-label *assortment* exists in each department.
   That is the main rival explanation for the pattern.

## Definition of "buys private label"

`BRAND == "Private"` in `product.csv`; the behavioural measure is **share of
dollars**, not items or trips:

- Merchandising decisions and private-label margin both accrue on dollars.
- Share of **items** over-weights cheap, high-count products.
- Share of **trips containing one** saturates near 1 and cannot discriminate.

A per-item sensitivity check appears in Section 8.

---

# Packages

```{r libs}
library(data.table)   # fast read of the 2.6M-row transaction file
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(scales)
library(knitr)
library(kableExtra)
library(tibble)
library(ggrepel)

theme_set(
  theme_minimal(base_size = 12) +
    theme(
      panel.grid.minor = element_blank(),
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      legend.position = "top"
    )
)
pal <- c(Private = "#00529B", National = "#BDC3C7")   # Kroger blue / grey
dir.create("figs", showWarnings = FALSE)
dir.create("out",  showWarnings = FALSE)
```

# Load the three datasets

The files live at the absolute paths given in `params`. `fread()` auto-detects
the delimiter. **`integer64 = "character"`** matters here: `BASKET_ID` values
such as `26984851472` exceed 32-bit integer range, and reading them as doubles
risks silent precision loss — they are identifiers, never arithmetic operands,
so character is the correct storage type.

```{r load}
trans <- fread(params$trans_path, integer64 = "character",
               showProgress = FALSE)
prod  <- fread(params$prod_path,  integer64 = "character",
               showProgress = FALSE)
demo  <- fread(params$demo_path,  integer64 = "character",
               showProgress = FALSE)

# household_key is lower-case in both transaction and demographic files;
# everything else is upper-case. Normalise so the join key is unambiguous.
setnames(trans, toupper(names(trans)))
setnames(prod,  toupper(names(prod)))
setnames(demo,  toupper(names(demo)))

stopifnot(all(c("HOUSEHOLD_KEY", "BASKET_ID", "PRODUCT_ID",
                "SALES_VALUE", "QUANTITY") %in% names(trans)))
stopifnot(all(c("PRODUCT_ID", "BRAND", "DEPARTMENT",
                "COMMODITY_DESC", "SUB_COMMODITY_DESC") %in% names(prod)))

cat("transaction_data:", format(nrow(trans), big.mark = ","), "rows\n")
cat("product:         ", format(nrow(prod),  big.mark = ","), "rows\n")
cat("hh_demographic:  ", format(nrow(demo),  big.mark = ","), "rows\n")
cat("\nColumns present:\n")
cat(" trans:", paste(names(trans), collapse = ", "), "\n")
cat(" prod :", paste(names(prod),  collapse = ", "), "\n")
cat(" demo :", paste(names(demo),  collapse = ", "), "\n")
```

## Cleaning decisions (each one deliberate and disclosed)

```{r clean}
# --- product file: character fields are space-padded in this dataset ---
chr_cols <- names(prod)[sapply(prod, is.character)]
prod[, (chr_cols) := lapply(.SD, str_squish), .SDcols = chr_cols]

prod[, BRAND               := str_to_title(BRAND)]
prod[, DEPARTMENT          := str_to_upper(DEPARTMENT)]
prod[, COMMODITY_DESC      := str_to_upper(COMMODITY_DESC)]
prod[, SUB_COMMODITY_DESC  := str_to_upper(SUB_COMMODITY_DESC)]
prod[DEPARTMENT == "" | is.na(DEPARTMENT), DEPARTMENT := NA_character_]

# How is BRAND actually coded? Never assume. If this table does not show
# "Private" and "National", the shares below will be wrong.
prod[, .N, by = BRAND][order(-N)]
```

```{r clean2}
# Keep only products whose brand AND department are known: an unlabelled brand
# cannot be scored either way, and including it would silently deflate shares.
prod_keep <- prod[BRAND %in% c("Private", "National") & !is.na(DEPARTMENT),
                  .(PRODUCT_ID, BRAND, DEPARTMENT,
                    COMMODITY_DESC, SUB_COMMODITY_DESC)]

# --- transactions: keep genuine sales only ---
# SALES_VALUE is dollars actually paid AFTER retail/loyalty discount. That is
# the money Kroger received, so no discount column is added back.
n0 <- nrow(trans)
tx <- trans[SALES_VALUE > 0 & QUANTITY > 0,
            .(HOUSEHOLD_KEY, BASKET_ID, PRODUCT_ID,
              SALES_VALUE, QUANTITY, STORE_ID, WEEK_NO)]
```

```{r clean3}
# --- join line items to product attributes (inner join drops unknown brand) ---
dt <- merge(tx, prod_keep, by = "PRODUCT_ID", all.x = FALSE)

audit <- tibble::tibble(
  step = c("raw transaction rows",
           "after dropping SALES_VALUE<=0 or QUANTITY<=0",
           "after inner join to products with known BRAND + DEPARTMENT"),
  rows = c(n0, nrow(tx), nrow(dt)),
  dollars = c(sum(trans$SALES_VALUE), sum(tx$SALES_VALUE), sum(dt$SALES_VALUE))
) %>%
  mutate(pct_rows_kept    = rows / first(rows),
         pct_dollars_kept = dollars / first(dollars))

audit %>%
  mutate(rows = comma(rows), dollars = dollar(dollars, accuracy = 1),
         across(starts_with("pct"), ~percent(.x, 0.1))) %>%
  kable(caption = "Data-handling audit trail: what was dropped and what it cost",
        col.names = c("Step", "Rows", "Dollars", "% rows kept", "% dollars kept")) %>%
  kable_styling(full_width = FALSE)
```

> **Read this table before believing anything downstream.** If the join retains
> well above 90% of dollars, the exclusions are a rounding error. If not, the
> excluded rows must be characterised before any share is quoted.

---

# Description first: what the store looks like

## Every department in the data

```{r dept-inventory}
dept_tbl <- dt[, .(
  revenue_total   = sum(SALES_VALUE),
  revenue_private = sum(SALES_VALUE[BRAND == "Private"]),
  items_total     = sum(QUANTITY),
  items_private   = sum(QUANTITY[BRAND == "Private"]),
  line_items      = .N,
  households      = uniqueN(HOUSEHOLD_KEY),
  products        = uniqueN(PRODUCT_ID)
), by = DEPARTMENT]

# assortment: how many DISTINCT private-label products exist per department
assort <- prod_keep[, .(products_all     = uniqueN(PRODUCT_ID),
                        products_private = uniqueN(PRODUCT_ID[BRAND == "Private"])),
                    by = DEPARTMENT]

dept_tbl <- merge(dept_tbl, assort, by = "DEPARTMENT", all.x = TRUE)
dept_tbl[, `:=`(
  pl_dollar_share = revenue_private / revenue_total,
  pl_item_share   = items_private   / items_total,
  pl_assort_share = products_private / products_all,
  rev_share_of_store = revenue_total / sum(revenue_total)
)]
setorder(dept_tbl, -revenue_private)

fwrite(dept_tbl, "out/table_department_summary.csv")

dept_tbl %>%
  as_tibble() %>%
  transmute(Department = DEPARTMENT,
            `Dept revenue` = dollar(revenue_total, accuracy = 1),
            `PL revenue`   = dollar(revenue_private, accuracy = 1),
            `PL $ share`   = percent(pl_dollar_share, 0.1),
            `PL item share`= percent(pl_item_share, 0.1),
            `PL assortment`= percent(pl_assort_share, 0.1),
            Households     = comma(households)) %>%
  kable(caption = "All departments, ranked by private-label revenue") %>%
  kable_styling(full_width = FALSE, font_size = 11) %>%
  scroll_box(height = "430px")
```

## The two rankings disagree — that is the finding

```{r ranking-clash}
rank_cmp <- dept_tbl[, .(DEPARTMENT, revenue_private, pl_dollar_share)][
  , `:=`(rank_by_revenue = frank(-revenue_private),
         rank_by_share   = frank(-pl_dollar_share))]

sp <- suppressWarnings(
  cor(rank_cmp$revenue_private, rank_cmp$pl_dollar_share, method = "spearman"))

top_rev   <- rank_cmp[which.min(rank_by_revenue)]
top_share <- rank_cmp[which.min(rank_by_share)]

cat("Top department by PL REVENUE :", top_rev$DEPARTMENT,
    "-", dollar(top_rev$revenue_private), "\n")
cat("Top department by PL SHARE   :", top_share$DEPARTMENT,
    "-", percent(top_share$pl_dollar_share, 0.1), "\n")
cat("Spearman correlation between the two rankings:", round(sp, 3), "\n")
```

```{r fig1, fig.cap="Slide A: private-label revenue and private-label share rank departments differently."}
plot_dat <- dept_tbl[order(-revenue_total)][1:min(12, .N)]
plot_dat[, DEPARTMENT := factor(DEPARTMENT, levels = rev(DEPARTMENT))]

long <- melt(plot_dat[, .(DEPARTMENT,
                          Private  = revenue_private,
                          National = revenue_total - revenue_private)],
             id.vars = "DEPARTMENT", variable.name = "BRAND",
             value.name = "revenue")

fig1 <- ggplot(long, aes(revenue, DEPARTMENT, fill = BRAND)) +
  geom_col(width = 0.7) +
  geom_text(data = plot_dat,
            aes(x = revenue_total, y = DEPARTMENT,
                label = percent(pl_dollar_share, 0.1)),
            inherit.aes = FALSE, hjust = -0.12, size = 3.3, colour = "#00529B") +
  scale_fill_manual(values = pal, name = NULL) +
  scale_x_continuous(labels = dollar_format(scale = 1e-3, suffix = "K"),
                     expand = expansion(mult = c(0, 0.16))) +
  labs(title = "Where the private-label business is, and where it wins",
       subtitle = "Bars: two-year revenue split. Blue label: private-label share of that department's dollars.",
       x = "Revenue (thousands of dollars)", y = NULL)
fig1
ggsave("figs/fig1_revenue_split.png", fig1, width = 9, height = 5.4, dpi = 300)
```

## Category level: what a category manager actually owns

`DEPARTMENT` is too coarse for a shelf decision, so the same ranking is run on
`DEPARTMENT + COMMODITY_DESC`. Both keys are used because identical commodity
names can occur in different departments.

```{r category}
cat_tbl <- dt[, .(revenue_total   = sum(SALES_VALUE),
                  revenue_private = sum(SALES_VALUE[BRAND == "Private"]),
                  households      = uniqueN(HOUSEHOLD_KEY)),
              by = .(DEPARTMENT, COMMODITY_DESC)]
cat_tbl[, pl_dollar_share := revenue_private / revenue_total]
setorder(cat_tbl, -revenue_private)
fwrite(cat_tbl, "out/table_category_summary.csv")

top_cat <- cat_tbl[1]

cat_tbl[1:15] %>% as_tibble() %>%
  transmute(Department = DEPARTMENT, Category = COMMODITY_DESC,
            `PL revenue` = dollar(revenue_private, accuracy = 1),
            `Category revenue` = dollar(revenue_total, accuracy = 1),
            `PL $ share` = percent(pl_dollar_share, 0.1),
            Households = comma(households)) %>%
  kable(caption = "Top 15 categories by private-label revenue") %>%
  kable_styling(full_width = FALSE, font_size = 11)
```

---

# Switching units: household x department

Ranking totals needs no test. But the moment a claim becomes *"private label
performs better in A than in B,"* the line item is the wrong unit. All
inference below uses **one household's private-label dollar share within one
department**.

A household is excluded from a department if it spent under
`r dollar(params$min_hh_dept)` there across two years: a $4 spend produces a 0%
or 100% share by accident, and those artefacts would inflate the variance and
distort every mean.

```{r hh-dept}
hh_dept <- dt[, .(spend_total   = sum(SALES_VALUE),
                  spend_private = sum(SALES_VALUE[BRAND == "Private"])),
              by = .(HOUSEHOLD_KEY, DEPARTMENT)]
hh_dept[, pl_share := spend_private / spend_total]

n_before <- nrow(hh_dept)
hh_dept  <- hh_dept[spend_total >= params$min_hh_dept]

cat("household x department cells before screen:", comma(n_before), "\n")
cat("after the >= $", params$min_hh_dept, " screen:      ",
    comma(nrow(hh_dept)), "\n", sep = "")
cat("distinct households retained:", comma(uniqueN(hh_dept$HOUSEHOLD_KEY)), "\n")
```

## Choosing the focus departments

The `r params$n_focus` departments carried into the tests are chosen on
**business relevance, before looking at any test result**: the departments with
the largest private-label revenue that also clear a minimum household count.
Choosing them by the size of the gap would be selecting on the outcome.

```{r focus}
elig <- hh_dept[, .(n_hh = .N), by = DEPARTMENT][n_hh >= 200]
focus_depts <- dept_tbl[DEPARTMENT %in% elig$DEPARTMENT][
  order(-revenue_private)][1:params$n_focus, DEPARTMENT]

cat("Focus departments:\n"); print(focus_depts)

hd <- hh_dept[DEPARTMENT %in% focus_depts]
hd[, DEPARTMENT := factor(DEPARTMENT, levels = focus_depts)]
```

## Distribution of the outcome, before any test

```{r fig2, fig.cap="Household-level private-label share is bounded, skewed and piled near zero in some departments. This is why a rank-based confirmation is reported alongside the ANOVA."}
fig2 <- ggplot(hd, aes(pl_share, DEPARTMENT)) +
  geom_violin(fill = "#00529B", alpha = 0.18, colour = NA, scale = "width") +
  geom_boxplot(width = 0.18, outlier.size = 0.4, outlier.alpha = 0.25) +
  stat_summary(fun = mean, geom = "point", shape = 23, size = 2.6,
               fill = "white", colour = "#00529B") +
  scale_x_continuous(labels = percent_format(1)) +
  labs(title = "Distribution of household private-label share, by department",
       subtitle = "One observation = one household's share of dollars in that department. Diamond = mean.",
       x = "Private-label share of household's spend in the department", y = NULL)
fig2
ggsave("figs/fig2_distribution.png", fig2, width = 9, height = 5.4, dpi = 300)
```

---

# Uncertainty: how big, and how sure

## Cluster bootstrap over households

A household appears in several departments, so its observations are not
independent. Resampling **households** (not rows) preserves that dependence; a
textbook t-interval ignores it and comes out too narrow. Both are reported so
the difference is visible.

```{r boot}
hh_ids <- unique(hd$HOUSEHOLD_KEY)
setkey(hd, HOUSEHOLD_KEY)

boot_mat <- matrix(NA_real_, nrow = params$boot_reps, ncol = length(focus_depts),
                   dimnames = list(NULL, as.character(focus_depts)))

for (b in seq_len(params$boot_reps)) {
  samp <- hd[.(sample(hh_ids, length(hh_ids), replace = TRUE)), allow.cartesian = TRUE]
  m <- samp[, .(m = mean(pl_share)), by = DEPARTMENT]
  boot_mat[b, as.character(m$DEPARTMENT)] <- m$m
}

ci_tbl <- hd[, .(n_hh = .N,
                 mean_share = mean(pl_share),
                 sd_share   = sd(pl_share),
                 median_share = median(pl_share)), by = DEPARTMENT]
ci_tbl[, `:=`(t_lo = mean_share - qt(0.975, n_hh - 1) * sd_share / sqrt(n_hh),
              t_hi = mean_share + qt(0.975, n_hh - 1) * sd_share / sqrt(n_hh))]
ci_tbl[, `:=`(
  boot_lo = apply(boot_mat[, as.character(DEPARTMENT), drop = FALSE], 2,
                  quantile, 0.025, na.rm = TRUE),
  boot_hi = apply(boot_mat[, as.character(DEPARTMENT), drop = FALSE], 2,
                  quantile, 0.975, na.rm = TRUE))]
setorder(ci_tbl, -mean_share)
fwrite(ci_tbl, "out/table_department_CIs.csv")

ci_tbl %>% as_tibble() %>%
  transmute(Department = DEPARTMENT, Households = comma(n_hh),
            `Mean PL share` = percent(mean_share, 0.1),
            `Median` = percent(median_share, 0.1),
            `95% CI (t)` = paste0(percent(t_lo, 0.1), " – ", percent(t_hi, 0.1)),
            `95% CI (cluster bootstrap)` =
              paste0(percent(boot_lo, 0.1), " – ", percent(boot_hi, 0.1))) %>%
  kable(caption = paste0("Mean household private-label share with 95% intervals (",
                         params$boot_reps, " bootstrap replications)")) %>%
  kable_styling(full_width = FALSE)
```

```{r fig3, fig.cap="Slide B: mean household private-label share with cluster-bootstrap 95% intervals. Non-overlapping intervals separate the departments."}
ci_plot <- copy(ci_tbl)[, DEPARTMENT := factor(DEPARTMENT, levels = rev(DEPARTMENT))]

fig3 <- ggplot(ci_plot, aes(mean_share, DEPARTMENT)) +
  geom_errorbarh(aes(xmin = boot_lo, xmax = boot_hi), height = 0.16,
                 colour = "#00529B", linewidth = 0.8) +
  geom_point(size = 3.2, colour = "#00529B") +
  geom_text(aes(label = percent(mean_share, 0.1)), vjust = -1.1, size = 3.4) +
  scale_x_continuous(labels = percent_format(1)) +
  labs(title = "How much of a household's department spend goes to private label",
       subtitle = paste0("Point = mean across households; bar = 95% cluster-bootstrap interval. ",
                         "Households with < $", params$min_hh_dept,
                         " in a department are excluded."),
       x = "Mean private-label share of household spend", y = NULL)
fig3
ggsave("figs/fig3_ci.png", fig3, width = 9, height = 5.4, dpi = 300)
```

## Is the overall difference real, and how large?

With tens of thousands of observations every p-value collapses to zero, so the
**effect size is the evidence** and the p-value is a formality.

```{r tests}
welch <- oneway.test(pl_share ~ DEPARTMENT, data = hd, var.equal = FALSE)
kw    <- kruskal.test(pl_share ~ DEPARTMENT, data = hd)

fit    <- aov(pl_share ~ DEPARTMENT, data = hd)
ss     <- summary(fit)[[1]][["Sum Sq"]]
eta2   <- ss[1] / sum(ss)
eps2   <- (unname(kw$statistic) - (length(focus_depts) - 1)) /
          (nrow(hd) - length(focus_depts))

lev <- car::leveneTest(pl_share ~ DEPARTMENT, data = hd)

cat("Welch one-way ANOVA: F =", round(welch$statistic, 1),
    ", p =", format.pval(welch$p.value, eps = .Machine$double.eps), "\n")
cat("Kruskal-Wallis:      chi2 =", round(kw$statistic, 1),
    ", p =", format.pval(kw$p.value, eps = .Machine$double.eps), "\n")
cat("Eta-squared (department share of variation): ",
    percent(eta2, 0.1), "\n", sep = "")
cat("Epsilon-squared (rank-based analogue):       ",
    percent(eps2, 0.1), "\n", sep = "")
cat("Levene test for equal variances: p =",
    format.pval(lev$`Pr(>F)`[1], eps = .Machine$double.eps),
    "-> Welch, not classical F\n")
```

Welch is used because the department standard deviations differ materially;
Kruskal–Wallis is reported alongside because the outcome is bounded in [0,1]
and skewed, and it reaches the same conclusion without any normality
assumption.

## Which specific departments differ — with multiplicity handled

`r params$n_focus` departments give
`r choose(params$n_focus, 2)` pairwise comparisons. Every one was run, and the
p-values are adjusted for all of them.

```{r pairwise}
tuk <- TukeyHSD(fit)$DEPARTMENT %>% as.data.frame() %>%
  rownames_to_column("comparison") %>% as_tibble() %>%
  arrange(desc(abs(diff)))

wil <- pairwise.wilcox.test(hd$pl_share, hd$DEPARTMENT,
                            p.adjust.method = "bonferroni")

tuk_out <- tuk %>%
  transmute(Comparison = comparison,
            `Gap (pp)` = round(diff * 100, 2),
            `95% CI (pp)` = paste0(round(lwr * 100, 2), " – ", round(upr * 100, 2)),
            `p (Tukey-adjusted)` = format.pval(`p adj`, digits = 3, eps = 1e-10),
            Verdict = ifelse(`p adj` < 0.05, "separated", "not separated"))
fwrite(tuk_out, "out/table_pairwise_tukey.csv")

tuk_out %>%
  kable(caption = paste0("All ", nrow(tuk),
                         " pairwise department comparisons, Tukey-adjusted. ",
                         "Gaps in percentage points of private-label share.")) %>%
  kable_styling(full_width = FALSE, font_size = 11)

n_sig <- sum(tuk$`p adj` < 0.05)
cat("Comparisons run:", nrow(tuk),
    "| surviving Tukey adjustment at 5%:", n_sig, "\n")
cat("Bonferroni-adjusted Wilcoxon agrees on:",
    sum(wil$p.value < 0.05, na.rm = TRUE), "of",
    sum(!is.na(wil$p.value)), "comparisons\n")
```

---

# Interrogating the finding

## Alternative explanation 1 — is it just shelf space?

If a department simply *stocks* more private-label items, a higher share is
arithmetic, not preference. The **preference index** divides the dollar share
by the assortment share: above 1 means shoppers choose private label more often
than its shelf presence would predict.

```{r fig4, fig.cap="Assortment versus performance. Departments above the diagonal convert shelf presence into dollars better than average."}
pref <- dept_tbl[DEPARTMENT %in% focus_depts,
                 .(DEPARTMENT, pl_assort_share, pl_dollar_share,
                   preference_index = pl_dollar_share / pl_assort_share)]

pref %>% as_tibble() %>% arrange(desc(preference_index)) %>%
  transmute(Department = DEPARTMENT,
            `PL share of assortment` = percent(pl_assort_share, 0.1),
            `PL share of dollars`    = percent(pl_dollar_share, 0.1),
            `Preference index`       = round(preference_index, 2)) %>%
  kable(caption = "Preference index = dollar share / assortment share. >1 means shoppers over-select private label relative to how much of it is on the shelf.") %>%
  kable_styling(full_width = FALSE)

fig4 <- ggplot(pref, aes(pl_assort_share, pl_dollar_share)) +
  geom_abline(slope = 1, intercept = 0, linetype = 2, colour = "grey55") +
  geom_point(size = 3.4, colour = "#00529B") +
  geom_text_repel(aes(label = DEPARTMENT), size = 3.3) +
  scale_x_continuous(labels = percent_format(1)) +
  scale_y_continuous(labels = percent_format(1)) +
  labs(title = "Shelf space does not fully explain the pattern",
       subtitle = "Dashed line: dollar share equals assortment share. Above it = genuine shopper preference.",
       x = "Private-label share of products stocked",
       y = "Private-label share of dollars")
fig4
ggsave("figs/fig4_assortment.png", fig4, width = 8, height = 5.4, dpi = 300)
```

## Alternative explanation 2 — is it just a different mix of shoppers?

Perhaps the departments differ because different *kinds* of households shop
them. Restricting to households that shop **all** focus departments holds
shopper composition constant by construction, and the Friedman test compares
the same households across departments.

```{r friedman}
complete_hh <- hd[, .N, by = HOUSEHOLD_KEY][N == length(focus_depts), HOUSEHOLD_KEY]
hd_c <- hd[HOUSEHOLD_KEY %in% complete_hh]

cat("Households shopping all", length(focus_depts), "focus departments:",
    comma(length(complete_hh)), "\n")

wide <- dcast(hd_c, HOUSEHOLD_KEY ~ DEPARTMENT, value.var = "pl_share")
mat  <- as.matrix(wide[, -1])
fr   <- friedman.test(mat)

kendall_w <- unname(fr$statistic) /
             (nrow(mat) * (ncol(mat) - 1))

cat("Friedman chi2 =", round(fr$statistic, 1),
    ", p =", format.pval(fr$p.value, eps = .Machine$double.eps),
    ", Kendall's W =", round(kendall_w, 3), "\n")

within_hh <- hd_c[, .(mean_share = mean(pl_share)), by = DEPARTMENT][order(-mean_share)]
within_hh %>% as_tibble() %>%
  transmute(Department = DEPARTMENT, `Mean PL share (same households)` =
              percent(mean_share, 0.1)) %>%
  kable(caption = "Within-household comparison: the ranking holds when shopper mix is held constant.") %>%
  kable_styling(full_width = FALSE)
```

The department ordering survives, so the pattern is not an artefact of which
households shop where.

## Independent cross-check on a different unit and metric

```{r chisq}
item_tab <- xtabs(QUANTITY ~ BRAND + DEPARTMENT,
                  data = dt[DEPARTMENT %in% focus_depts])
cs <- chisq.test(item_tab)
cramers_v <- sqrt(unname(cs$statistic) /
                    (sum(item_tab) * (min(dim(item_tab)) - 1)))

cat("Chi-square of association (items, line-item unit): X2 =",
    round(cs$statistic, 0),
    ", df =", cs$parameter,
    ", p =", format.pval(cs$p.value, eps = .Machine$double.eps), "\n")
cat("Cramer's V =", round(cramers_v, 3), "\n")
```

This uses a different unit (items, not households) and a different metric
(counts, not dollars). Agreement means the conclusion is not an artefact of
either choice. Note the caveat: at the item level the observations are not
independent, which is exactly why this is a cross-check and not the primary
test.

## Sensitivity: dollars versus items

```{r sensitivity}
dept_tbl[DEPARTMENT %in% focus_depts,
         .(DEPARTMENT, pl_dollar_share, pl_item_share)][
  , gap_pp := (pl_item_share - pl_dollar_share) * 100][
  order(-pl_dollar_share)] %>%
  as_tibble() %>%
  transmute(Department = DEPARTMENT,
            `PL share of dollars` = percent(pl_dollar_share, 0.1),
            `PL share of items`   = percent(pl_item_share, 0.1),
            `Item minus dollar (pp)` = round(gap_pp, 1)) %>%
  kable(caption = "Item share runs above dollar share wherever private label is cheaper per unit. The ranking is what matters — check whether it changes.") %>%
  kable_styling(full_width = FALSE)
```

---

# Limits of this evidence

- **Observational, not causal.** These are shares under whatever assortment,
  pricing and promotion were live at the time. Nothing here says that expanding
  private label in a department *would* raise its share. A controlled
  assortment or price test across matched stores is what would license that
  claim.
- **Brand is a coarse flag.** `BRAND` distinguishes private from national but
  not premium private label from value private label, which is a margin-relevant
  distinction this analysis cannot make.
- **No margin data.** The recommendation is framed in revenue and share because
  the files contain no cost. Whether a share point in one department is worth
  more than in another depends on margins we do not observe.
- **Two years, one retailer, unknown geography.** Store location and local
  competition are absent, so a department's performance may reflect the
  competitive set rather than shopper preference.
- **Screened households.** The `r dollar(params$min_hh_dept)` minimum makes the
  shares stable but tilts the sample toward committed shoppers of each
  department. Occasional shoppers are under-represented by design.

---

# Every number quoted on the slides

```{r slide-numbers}
slide <- c(
  paste0("departments_analysed              : ", nrow(dept_tbl)),
  paste0("departments_passing_screen        : ", nrow(elig)),
  paste0("focus_departments                 : ", paste(focus_depts, collapse = " | ")),
  paste0("households_total                  : ", comma(uniqueN(dt$HOUSEHOLD_KEY))),
  paste0("households_in_tests               : ", comma(uniqueN(hd$HOUSEHOLD_KEY))),
  paste0("hh_dept_observations              : ", comma(nrow(hd))),
  paste0("dollars_retained_after_cleaning   : ",
         percent(audit$pct_dollars_kept[3], 0.1)),
  "",
  paste0("top_dept_by_PL_revenue            : ", top_rev$DEPARTMENT),
  paste0("top_dept_PL_revenue               : ", dollar(top_rev$revenue_private)),
  paste0("top_dept_by_PL_share              : ", top_share$DEPARTMENT),
  paste0("top_dept_PL_share                 : ", percent(top_share$pl_dollar_share, 0.1)),
  paste0("spearman_revenue_vs_share         : ", round(sp, 3)),
  paste0("top_category_by_PL_revenue        : ", top_cat$DEPARTMENT, " / ",
         top_cat$COMMODITY_DESC),
  paste0("top_category_PL_revenue           : ", dollar(top_cat$revenue_private)),
  paste0("top_category_PL_share             : ", percent(top_cat$pl_dollar_share, 0.1)),
  "",
  paste0("highest_mean_hh_share_dept        : ", ci_tbl$DEPARTMENT[1]),
  paste0("highest_mean_hh_share             : ", percent(ci_tbl$mean_share[1], 0.1),
         "  CI ", percent(ci_tbl$boot_lo[1], 0.1), " - ",
         percent(ci_tbl$boot_hi[1], 0.1)),
  paste0("lowest_mean_hh_share_dept         : ", ci_tbl$DEPARTMENT[nrow(ci_tbl)]),
  paste0("lowest_mean_hh_share              : ",
         percent(ci_tbl$mean_share[nrow(ci_tbl)], 0.1),
         "  CI ", percent(ci_tbl$boot_lo[nrow(ci_tbl)], 0.1), " - ",
         percent(ci_tbl$boot_hi[nrow(ci_tbl)], 0.1)),
  paste0("top_minus_bottom_gap_pp           : ",
         round((ci_tbl$mean_share[1] - ci_tbl$mean_share[nrow(ci_tbl)]) * 100, 1)),
  "",
  paste0("welch_F                           : ", round(welch$statistic, 1)),
  paste0("welch_p                           : ",
         format.pval(welch$p.value, eps = .Machine$double.eps)),
  paste0("kruskal_p                         : ",
         format.pval(kw$p.value, eps = .Machine$double.eps)),
  paste0("eta_squared                       : ", percent(eta2, 0.1)),
  paste0("pairwise_comparisons_run          : ", nrow(tuk)),
  paste0("pairwise_significant_adjusted     : ", n_sig),
  paste0("largest_gap_comparison            : ", tuk$comparison[1],
         " = ", round(tuk$diff[1] * 100, 1), " pp  CI [",
         round(tuk$lwr[1] * 100, 1), ", ", round(tuk$upr[1] * 100, 1), "]"),
  "",
  paste0("friedman_p                        : ",
         format.pval(fr$p.value, eps = .Machine$double.eps)),
  paste0("kendalls_W                        : ", round(kendall_w, 3)),
  paste0("complete_households               : ", comma(length(complete_hh))),
  paste0("cramers_V_items                   : ", round(cramers_v, 3)),
  paste0("highest_preference_index          : ",
         pref[which.max(preference_index), DEPARTMENT], " = ",
         round(max(pref$preference_index), 2))
)
writeLines(slide, "out/slide_numbers.txt")
cat(slide, sep = "\n")
```

```{r session, echo=FALSE}
sessionInfo()
```
